Cloud correction and filtering of operational microwave humidity radiances with case specific uncertainty estimation

### 2020-09-11 Article/ici_submm_cloudcorr.tex
A first complete, but raw, version of the Introduction.

%% 2-column papers and preprints
\documentclass[amt, manuscript]{copernicus}

@@

\title{An emerging approach for cloud correction and filtering of operational
  microwave humidity radiances}

@@

\runningtitle{A cloud correction approach}

@@

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitions imposed by clouds. To be able to also sense humidity
inside and below clouds weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

Precipiation and most dense clouds, particularly if found at high altitude, can
still afffect measured radiances around 183\,GHz
\citep[e.g.][]{bennartz2003sensitivity}. As the impact from the hydrometeors
then is dominated by scattering, the complexity of the analysis of the data
increases dramatically and there exists a need to identify the problematic
cases. This is normally denoted as cloud filtering, to obtain data of ``clear
sky'' character. Such filtering has been applied to derive climate records
\citep{lang2020new} and is essentail in studies of the agreement between
observations and simulations \citep{brogniez2016review} as well as 
comparing observations of different instruments to validate their calibration
\citep{john2013assessment,moradi:retri:15,berg2016intercalibration}. A commonly
used cloud filtering methods for these applicatons is the one of
\citet{buehler:aclou:07}. This method is based on the 183\,GHz data alone,
involving rules on the brightness temperatures differences between channels. An
older version is \citet{burns1997effects}.

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of microwave data by ``all-sky''
assimilation is growing \citep{geer2017growing}, but 183\,GHz data are still
mainly used in a clear sky fashion \citep{geer2018all}. The later is
particularly true in NWP of regional scope, all-sky assimilation of 183\,GHz at
many national weather agencies will likely not be reached in many years. The
cloud filtering in NWP differs between centres. One approach is to make use of
a ``scattering index'' \citep{bennartz2002precipitation}, based on the
observations alone. Today, likely more common is to use ``observation minus
background'' (O-B), where the forecast model is used to obtain an estimate of
the expected clear-sky value and the observation is rejected if the deviation
exceeds some threshold (???). In both cases, the filtering typically involving
observations around 89 and/or 150\,GHz.

Despite the broad usage these filtering methods have some drawbacks. For
183\,GHz, the impact of hydrometeors systematically cause a decrease in the
observed radiance (maybe with exceptions at extremely dry conditions), see
e.g.\ \citet{barlakas:three:20} and below. This means that if any cloud
contaminated is missed by the filtering this will cause a negative bias in mean
radiance, compared to the true clear-sky mean. This bias will translate to
a bias in humidity after the retrieval or assimilation. The alternative is to
apply a very strict filtering, but this will result in that a high fraction of
actually clear-sky values will be rejected \todo{(Reka or better reference)},
i.e.\ an important loss of useful data. One for all. Same or lower frequency ...

% favrichon2019detecting?

>>>>

%% 2-column papers and preprints
\documentclass[amt]{copernicus}

@@

\title{Cloud correction and filtering of operational microwave humidity
  radiances with case specific uncertainty estimation}

@@

\runningtitle{Cloud correction with case specific uncertainty estimation}

@@

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

Precipitation and most dense clouds, particularly if found at high altitude, can
still affect measured radiances around 183\,GHz
\citep[e.g.][]{bennartz2003sensitivity}. As the impact from the hydrometeors
then is dominated by scattering, the complexity of the analysis of the data
increases dramatically and there exists a need to identify the problematic
cases. This is normally denoted as cloud filtering, to obtain data of ``clear
sky'' character. Such filtering has been applied to derive climate records
\citep{lang2020new} and is essential in studies of the agreement between
observations and simulations \citep{brogniez2016review} as well as 
comparing observations of different instruments to validate their calibration
\citep{john2013assessment,moradi:retri:15,berg2016intercalibration}. A commonly
used cloud filtering methods for these applications is the one of
\citet{buehler:aclou:07}. This method is based on the 183\,GHz data alone,
involving rules on the brightness temperatures differences between channels. An
older version is \citet{burns1997effects}.

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of microwave data by ``all-sky''
assimilation is growing \citep{geer2017growing}, but 183\,GHz data are still
mainly used in a clear sky fashion \citep{geer2018all}. The later is
particularly true in NWP of regional scope, all-sky assimilation of 183\,GHz at
many national weather agencies will likely not be reached in many years. The
cloud filtering in NWP differs between centres. One approach is to make use of
a ``scattering index'' \citep{bennartz2002precipitation}, based on the
observations alone. Today, likely more common is to use ``observation minus
background'' (O-B), where the forecast model is used to obtain an estimate of
the expected clear-sky value and the observation is rejected if the deviation
exceeds some threshold (???). In both cases, the filtering typically involves
observations around 89 and/or 150\,GHz.

Despite the broad usage these filtering methods have some drawbacks. For
183\,GHz, the impact of hydrometeors systematically causes a decrease in the
observed radiance (maybe with exceptions at extremely dry conditions), see
e.g.\ \citet{barlakas:three:20} and below. This means that if any cloud
contamination is missed by the filtering this will cause a negative bias in
mean radiance, compared to the true clear-sky mean. This bias will translate to
a bias in humidity after the retrieval or assimilation. The alternative is to
apply a very strict filtering, but this will result in that a high fraction of
actually clear-sky values will be rejected \todo{(Reka or better reference)},
i.e.\ an important loss of useful data.

The filtering is normally done in a ``one for all'' manner, i.e.\ all 183\,GHz
channels are either kept or rejected, while, as the channels differ in their
altitude coverage, there could be cloud impact in some channels and still the
others can be considered as clear-sky. To allow a channel specific filtering,
data likely need to be combined in a more complex manner than simple
differences, but it is unclear what type of regression that would be best. This
points towards applying machine learning, as used by e.g.\
\citet{favrichon2019detecting}.

A maybe less obvious problem is the uncertainty to assign to the filtered
values. To our best knowledge, in best case estimates of mean and worst case
errors are provided. Some cases with relatively high cloud impact will likely be
missed, while most cases are clear sky from start. As the remaining cloudy
cases can cause significant biases, the likely solution is to apply a quite
conservative (high) error estimate. However, this will unnecessary downgrade
the value of the truly clear sky cases and the observations are used in a
non-optimal manner.

We are here approaching cloud filtering task from a new angle. The basic idea
is to derive an estimate of the corresponding noise-free clear-sky (NFCS) value
(i.e.\ the radiance that would have been measured in absence of noise and
hydrometeors). This is done for each channel separately, only using
measurements (no ``background'' data involved). Not only a best estimate is
provided, but also a case specific uncertainty.

This information can be used as a pure filter, by rejecting data where the
correction exceeds some threshold value. This threshold value should be
relatively low, to not introduce a bias in filtered dataset. However, even
better is to replace the original value with the predicted NFSC value when
forming the clear-sky dataset. This approach we denote as cloud correction. It
is shown below that a basically bias free cloud correction can be obtained.
This feature removes the need of threshold value, as long as the retrieval or
assimilation system can incorporate the uncertainty of the corrected value. As
also will be shown, the uncertainty for originally clear sky data is determined
by noise, but the uncertainty increases with magnitude of correction.
Accordingly, the cloud correction approach allows that the full weight of
clear-sky data can be preserved.

The estimation of NFSC values makes use of a special type of machine learning,
denoted as Quantile Regression Neural Network (QRNN,
\citet{pfreundschuh:aneur:18}). Unlike standard usage of machine learning where
only some kind of best estimate is provided, QRNN works in a Bayesian fashion
and instead gives a description of the posterior uncertainty. More precisely,
QRNN outputs an user specified set of percentiles that gives a discrete
description posterior distribution, that can be process to derive e.g.\ the
expectation value.

The approach is demonstrated on a combination of 183\,GHz channels (following
\citet{buehler:aclou:07}), but we mainly explore the cloud correction (of
183\,GHz data) that will be possible when sub-millimetre data will be at hand
in some years. This wavelength region will be introduced by the Ice Cloud
Imager (ICI, \citet{eriksson:towar:20}), but likely also be included in several
smaller missions (such as AWS presented below). The focus on ICI is motivated
by several reasons. The true potentail of the QRNN cloud correction approach is
likely not revealed if tested on existing data and applying ICI data for cloud
filtering has not earlier been discussed. It is also argued that the cloud
correction makes it possible to make good use of ICI data already in clear-sky
assimilation, a fact that should speed up the use of this novel data source.
